---
title: "Canada Basemap"
subtitle: "SUBTITLE"
author: "Spencer Zeigler"
date: "Last knitted `r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: console # switch to 'inline' if preferred
---

# Setup

```{r "setup", message = FALSE}
# packages
library(sf) # the base package manipulating shapes
library(rgdal) # geo data abstraction library
library(geojsonio) # geo json input and output
library(spdplyr) # the `dplyr` counterpart for shapes
library(rmapshaper) # the package that allows geo shape transformation
library(magrittr) # data wrangling
library(dplyr)
library(tidyr)
library(ggplot2) # general data wrangling and plotting
library(sp)

# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("plots", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

## SOURCES
https://tengl.net/blog/2020/1/7/drawing-canada-maps-in-r
https://r-spatial.org/r/2018/10/25/ggplot2-sf-2.html

# Convert map of Canada from .shp to geojson
```{r}
#canada_raw = readOGR(dsn = "data", layer = "canada_basemap", encoding = 'latin1') # 1
#canada_raw_json <- geojson_json(canada_raw) # 2
#canada_raw_sim <- ms_simplify(canada_raw_json) # 3
#geojson_write(canada_raw_sim, file = "data/canada_cd_sim.geojson") # 4
```

#Convert shapefile with approximate crop shape from .shp to geojson
```{r}
# crop_raw = readOGR(dsn = "data", layer = "CanadaCrop", encoding = 'latin1') 
# crop_raw_json <- geojson_json(crop_raw) 
# geojson_write(crop_raw_json, file = "data/crop_cd_sim.geojson") 
```

# load CRS; read in map and shapefile (as geojson); transform map to correc CRS
```{r}
crs_string <- "+proj=lcc +lat_1=49 +lat_2=77 +lon_0=-91.52 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs" # 2

canada_cd <- st_read("data/canada_cd_sim.geojson", quiet = TRUE) # canada
canada_cd_t <- st_transform(canada_cd, crs_string) # canada with transformed to correct crs

crop_cd <- st_read("data/crop_cd_sim.geojson", quiet = TRUE) # crop rectangle

st_crs(canada_cd_t) # check crs
#st_crs(crop_cd) <- crs_string # use this line only if the geometry go to 0 after using st_transform
```


# Aesthetics 
```{r}
# Define the maps' theme -- remove axes, ticks, borders, legends, etc.
theme_map <- function(base_size=9, base_family="") { # 3
	require(grid)
	theme_bw(base_size=base_size, base_family=base_family) %+replace%
		theme(axis.line=element_blank(),
			  axis.text=element_blank(),
			  axis.ticks=element_blank(),
			  axis.title=element_blank(),
			  panel.background=element_blank(),
			  panel.border=element_blank(),
			  panel.grid=element_blank(),
			  panel.spacing=unit(0, "lines"),
			  plot.background=element_blank(),
			  legend.justification = c(0,0),
			  legend.position = c(0,0)
		)
}

# Define the filling colors for each province; max allowed is 9 but good enough for the 13 provinces + territories
map_colors <- RColorBrewer::brewer.pal(9, "Pastel1") %>% rep(37) # 4
```

# Plotting Canada
```{r}
# Plot the map
ggplot() +
  geom_sf(data = canada_cd, color = "gray60", size = 0.1) +
  coord_sf(default_crs = NULL) +
	scale_fill_manual(values = map_colors) +
	guides(fill = "none") +
	theme_map() + 
  theme(panel.grid.major = element_line(color = "white"),
        legend.key = element_rect(color = "gray40", size = 0.1))
```

# Trying to figure out how to crop the map. 
```{r}
# QGIS plots my Canada data in NAD83/ Statistics Canada Lambert EPSG: 3357
#st_set_crs(canada_cd, 3347)

#Create bbox object to use with st_crop().
bbox <- st_bbox(crop_cd[[2]][[1]]) # turn xmin,xmax,ymin,ymax from crop into bbox
bbox <- st_set_crs(bbox, crs_string) # set crs of bbox 
st_crs(bbox) # check that the crs set

st_crop(canada_cd_t, bbox) #Error in FUN(X[[i]], ...) : !anyNA(x) is not TRUE
#c(-1585578.9, 4009743.7, 724394.2, 5221468.4)

st_is_valid(canada_cd_t) #checked this after getting !anyNA(x) not TRUE; saw that this produces some FALSE values
canada_cd_t <- st_make_valid(canada_cd_t) #attempt to fix FALSE values, it does. 
st_is_valid(canada_cd_t) #now they are all true. try running line 123 again and get the same !anyNA(x) error. 

```

# Since all cropping attemps have failed, here is a filtered dataset of the two provinces I care about. :)
# Not ideal since I still can't zoom in on specific areas I care about, but better than looking at all of Southern Canada.
```{r}
nwt_cd <- canada_cd_t %>%
  filter(PRNAME == "Northwest Territories / Territoires du Nord-Ouest" | PRNAME == "Nunavut") #%>% filter(!CDNAME == "Baffin")
```

# Kimberlite Data
```{r}
pipe_coords_raw <- readxl::read_excel("data/kimberlite_coords.xlsx") %>%
  drop_na(long) #read in data I want to plot on the map

pipe_coords <- st_as_sf(pipe_coords_raw, 
                     coords = c("long", "lat"), 
                     remove = FALSE, 
                     crs = 4269, 
                     agr = "constant") #transform pipe_coords_raw so they plot correctly on the map
```

#Plot with Kimberlite Data
```{r}
ggplot() +
  geom_sf(data = pipe_coords, aes(color = age_Ma), alpha = 1, size = 1) + # 
	geom_sf(data = nwt_cd, color = "gray60", fill = "lightblue", size = 0.1, alpha = 0.3) +
	coord_sf(default_crs = NULL) +
	guides(fill = "none") +
  theme_map() + 
  theme(panel.grid.major = element_line(color = "white"),
  legend.key = element_rect(color = "gray40", size = 0.1))
```












